Video-Agnostic Perturbations: A Simple and Strong Baseline for Siamese Visual Tracking Attack

### 2021-05-01 博士/论文提交/adversarial_attack/TCSVT/latex/imperceptible.tex
更新

\begin{equation}
\begin{gathered}
  \min\limits_{\delta_{\textbf{x}_{i}}, \delta_{\textbf{z}_{i}}} L(y^{fake}_i, f(\textbf{x}_i + \delta_{\textbf{x}_{i}}, \textbf{z}_0 + \delta_{\textbf{z}_{i}}))\\
  \text{s.t.}\ ||\delta_{\textbf{x}_i}||_p \le \epsilon, ||\delta_{\textbf{z}_i}||_p \le \epsilon,
\end{gathered}
\end{equation}
where $L$ is the loss function of the SiamFC++ tracker. $(\textbf{x}_i, \textbf{z}_0)$ is the search-template pair of the video being tracked. $y^{fake}_i=\{\textbf{C}^*, \textbf{R}^*, \textbf{Q}^*\}$ is the fake target of the loss function generated according to the fake ground truth box $b^{fake}_i$ that we want the tracker to output at frame $i$ (see Sec. \ref{generate}). 
% 不写 BIM，原因是BIM的优化为 xi+1 = xi + delta, 与上面公式不太好兼容。
However, these methods require perturbing the input image pair $(\textbf{x}, \textbf{z})$ frame by frame, which comes at a non-negligible cost, considering that VOT is a real-time task.
As a consequence, universal adversarial perturbations \cite{UAP, shafahi2020universal} are more practical for attacking Siamese trackers:
\begin{equation}
  \begin{gathered}
    \min\limits_{\delta_\textbf{x}, \delta_\textbf{z}} \mathop{\mathbb{E}}\limits_{(\textbf{x}, \textbf{z})} L(y^{fake}, f(\textbf{x} + \delta_\textbf{x}, \textbf{z} + \delta_\textbf{z}))\\
    \text{s.t.}\ ||\delta_\textbf{x}||_p \le \epsilon, ||\delta_\textbf{z}||_p \le \epsilon,
  \end{gathered}
\end{equation}  
where the input pair $(\textbf x,\textbf z)$ is randomly picked from the training set and useful for all testing set. 
Note that universal adversarial perterbations are added to the whole image and cannot tell the tracker where to look, so $y^{target}$ is kept constant during the training process, resulting in the inability to perform targted attacks.
To achieve universal targeted attack, one feasible way is to paste a small (universal) adversarial patch into the search images adhering to the predefined trajectory, so that the tracker outputs the location and size of the adversarial patch instead of the real target:
\begin{equation}
    \min\limits_{\delta_\textbf{x}} \mathop{\mathbb{E}}\limits_{(\textbf{x}, \textbf{z}, y^{fake})} L(y^{fake}, f(\textbf{x}\odot (1-m) + \delta_\textbf{x}\odot m, \textbf{z}))
\end{equation}
where $m$ is a binary mask and $m_{x,y}=1$ at all spatial locations under the fake ground bounding box $b^{fake}$. The target box is changed according to the patch position. $\odot$ is element-wise multiplication. Note that in this naive way to use patch, the template image is not changed. However, the patch is not constrained, which decrease the usage.
However, CNN attacks are usually expected imperceptible but the above method has to add an obviously noticeable fake target patch to tracking frames, which raises the risk of being suspected.

@@

>>>>

\begin{equation}
\begin{gathered}
  \min\limits_{\delta_{\textbf{x}_{i}}, \delta_{\textbf{z}_{i}}} L(y^{fake}_i, f(\textbf{x}_i + \delta_{\textbf{x}_{i}}, \textbf{z}_0 + \delta_{\textbf{z}_{i}}))\\
  \text{s.t.}\ ||\delta_{\textbf{x}_i}||_p \le \epsilon, ||\delta_{\textbf{z}_i}||_p \le \epsilon,
\end{gathered}
\end{equation}
where $L$ is the loss function of the SiamFC++ tracker. $(\textbf{x}_i, \textbf{z}_0)$ is the search-template pair of the video being tracked. $y^{fake}_i=\{\textbf{C}^*, \textbf{R}^*, \textbf{Q}^*\}$ is the fake target of the loss function generated according to the fake ground truth box $b^{fake}_i$ that we want the tracker to output at frame $i$ (see Sec. \ref{generate}). 
% 不写 BIM，原因是BIM的优化为 xi+1 = xi + delta, 与上面公式不太好兼容。
Additionally, $\delta$ should be sufficiently small, which is commonly modeled through an upper-bound $\epsilon$ on the $l_p\text{-norm}$, commonly denoted as $||\cdot||_p$, of the perturbation. A popular choice is to set $p=\inf$.
However, these methods require perturbing the input image pair $(\textbf{x}, \textbf{z})$ frame by frame, which comes at a non-negligible cost, considering that VOT is a real-time task.
As a consequence, universal adversarial perturbations \cite{UAP, shafahi2020universal} are more practical for attacking Siamese trackers:
\begin{equation}
  \begin{gathered}
    \min\limits_{\delta_\textbf{x}, \delta_\textbf{z}} \mathop{\mathbb{E}}\limits_{(\textbf{x}, \textbf{z})} L(y^{fake}, f(\textbf{x} + \delta_\textbf{x}, \textbf{z} + \delta_\textbf{z}))\\
    \text{s.t.}\ ||\delta_\textbf{x}||_p \le \epsilon, ||\delta_\textbf{z}||_p \le \epsilon,
  \end{gathered}
\end{equation}  
where the input pair $(\textbf x,\textbf z)$ is randomly picked from the training set and useful for all testing set. 
Note that universal adversarial perterbations are added to the whole image and cannot tell the tracker where to look, so $y^{target}$ is kept constant during the training process, resulting in the inability to perform targted attacks.
To achieve universal targeted attack, one feasible way is to paste a small (universal) adversarial patch into the search images adhering to the predefined trajectory, so that the tracker outputs the location and size of the adversarial patch instead of the real target:
\begin{equation}
    \min\limits_{\delta_\textbf{x}} \mathop{\mathbb{E}}\limits_{(\textbf{x}, \textbf{z}, y^{fake})} L(y^{fake}, f(\textbf{x}\odot (1-m) + \delta_\textbf{x}\odot m, \textbf{z}))
\end{equation}
where $m$ is a binary mask and $m_{x,y}=1$ at all spatial locations under the fake ground bounding box $b^{fake}$. The target box is changed according to the patch position. $\odot$ is element-wise multiplication. Note that in this naive way to use patch, the template image is not changed. However, the patch is not constrained, which decrease the usage.
However, CNN attacks are usually expected imperceptible but the above method has to paste an obviously noticeable fake target patch to tracking frames, which raises the risk of being suspected.

@@

\subsection{Compared with baselines}

We compare our attack method with two baseline methods (See Table \ref{tab:imperceptible}).
%\textbf{Baseline-1} uses the BIM \cite{DBLP:conf/iclr/KurakinGB17a} method to do targeted attack. The difference between the BIM method and our method includes: (1) We use the training set, while Baseline-1 use every image. (2) We perform imperceptible patch, while Baseline-1 use full image perturbations. It needs to be optimized on every frame. So it spped is slow.
\textbf{Baseline-1} performs untargeted attacks based on the UAP \cite{UAP} method. The difference between Baseline-1 and our method is that, Baseline-1 generates perturbations added to the whole search image and cannot perform targeted attacks, while our method adds a small (universal) imperceptible patch to the search image to perform targeted attacks (see Figure \ref{fig:imperceptible}).
\textbf{Baseline-2} performs targeted attacks based on the adversarial patch \cite{patch} method. The difference between Baseline-2 and our method includes: (1) Baseline-2 pastes a adversarial patch on the search image without the $l_p\text{-norm}$ constrain while our method adds the patch onto the search image with the $l_p\text{-norm}$ constrain. (2) Baseline-2 uses the clean template image while we add imperceptible perterbations onto the template image. As a consequence, Baseline-2 generates an obviously noticeable patch while ours are imperceptible as shown in Figure \ref{fig:imperceptible}. Moreover, both the untargeted attack performance and the targeted performance of our method are better than Baseline-2 as shown in Table \ref{tab:imperceptible}.
\begin{table}[t]
\centering
\caption{Compared with baselines on GOT-Val.}
\begin{tabular}{@{}lcc@{}}
  \toprule
  \multirow{2}{*}[-2pt]{Methods}  & \multicolumn{2}{c}{AO}\\ \cmidrule{2-3}
       & Real Traj. & Fake Traj.\\ \midrule
%Baseline-1 \cite{DBLP:conf/iclr/KurakinGB17a}  & ?          & ?\\
Baseline-1 \cite{UAP}  & 0.09          & -\\
Baseline-2 \cite{patch}   & 0.23       & 0.78\\
  Ours & 0.15       & 0.84\\ \bottomrule
  \end{tabular}
\label{tab:imperceptible}
\end{table}

\begin{figure}
  \centering
  %\subfigure[Baseline-1]{\label{fig:imperceptiblea}\includegraphics[width=0.11\textwidth]{images_imperceptible/imperceptible/AP.jpg}}
  \subfigure[Baseline-1]{\label{fig:imperceptibleb}\includegraphics[width=0.15\textwidth]{images_imperceptible/imperceptible/UAP.jpg}}
  \subfigure[Baseline-2]{\label{fig:imperceptiblec}\includegraphics[width=0.15\textwidth]{images_imperceptible/imperceptible/AP.jpg}}
  \subfigure[Ours]{\label{fig:imperceptibled}\includegraphics[width=0.15\textwidth]{images_imperceptible/imperceptible/Ours.jpg}}
  \caption{Visualization of perturbations on search images. Baseline-1 generates perturbations added to the whole search image and cannot perform targeted attacks. Baseline-2 pastes an obviously noticeable patch on the search image, which raises the risk of being suspected. Our method adds a small (universal) imperceptible patch to the search image to perform targeted attacks.}
  \label{fig:imperceptible}
\end{figure}
